---
title: "HIV_Data"
author: "Denglinsui"
date: "2021/3/16"
output:
  pdf_document: default
  html_document: default
bibliography: mybib.bib
---

```{r library, include=FALSE}
rm(list = ls())
# Output Table
library(knitr)

# Plot
library(ggplot2)
library(ggpubr)

# Tidy Dataset
library(dplyr)
library(reshape)

# Aprior Algorithm
library(arules)
library(arulesViz)

# To discover gene
library(knockoff)

setwd("~/project/HIV_Data/Comparison")
source("multilayer_knockoff_filter.R")
source("aux_mkf.R")
source("aux_knockoffs.R")
source("p_filter.R")


```


### Antiretroviral Drug
There are four major types of antiretroviral drugs: eight protease inhibitors(PIs), seven nucleoside and one nucleotide reverse transicriptase inhibitors(NRTIs), three nonnucleoside reverse transcriptase inhibitors(NNRTIs) amd one fusion inhibitor. The first three drug classes are included in this dataset.

```{r Drug_Class, echo=FALSE}
drug_class_all <- c("PI","NRTI","NNRTI")
Drug_extract <- function(drug_class){
  base_url = 'http://hivdb.stanford.edu/pages/published_analysis/genophenoPNAS2006'
  gene_url = paste(base_url, 'DATA', paste0(drug_class, '_DATA.txt'), sep='/')
  tsm_url = paste(base_url, 'MUTATIONLISTS', 'NP_TSM', drug_class, sep='/')
  
  gene_df = read.delim(gene_url, na.string = c('NA', ''), stringsAsFactors = FALSE)

  tsm_df = read.delim(tsm_url, header = FALSE, stringsAsFactors = FALSE)
  names(tsm_df) = c('Position', 'Mutations')

  pos_start = which(names(gene_df) == 'P1')
  Drug_name = paste(names(gene_df)[4:(pos_start-1)],collapse = " ")
  return(list(gene_df = gene_df,
              tsm_df = tsm_df,
              Drug_name = Drug_name,
              Drug_class = drug_class))
}

res_PI <- Drug_extract("PI")
res_NRTI <- Drug_extract("NRTI")
res_NNRTI <- Drug_extract("NNRTI")

Drug_summary <- data.frame(DrugClass = drug_class_all,
                           Drug = c(res_PI$Drug_name,
                                    res_NRTI$Drug_name,
                                    res_NNRTI$Drug_name))
row.names(Drug_summary) <- NULL
```


```{r GroundTruth}
PosMut = NULL
tsm_df <- res_PI$tsm_df
for(i in 1:nrow(tsm_df)){
  Pos <- tsm_df$Position[i]
  Mut <- strsplit(tsm_df$Mutations[i], split = ' ')[[1]]
  PosMut <- c(PosMut, paste0("P",Pos,".",Mut))
}
```




```{r knockoff_and_bhq, include=FALSE}
knockoff_and_bhq <- function (X, y, q) {
  # Log-transform the drug resistance measurements.
  y = log(y)
  
  # Remove patients with missing measurements.
  missing = is.na(y)
  y = y[!missing]
  X = X[!missing,]
  
  # Remove predictors that appear less than 3 times.
  X = X[,colSums(X) >= 3]
  
  # Remove duplicate predictors.
  X = X[,colSums(abs(cor(X)-1) < 1e-4) == 1]
  
  # Run the knockoff filter.
  knock.gen = function(x) create.fixed(x, method='equi')
  result = knockoff.filter(X, y, fdr=q, 
                           knockoffs=knock.gen, 
                           statistic=stat.glmnet_lambdasmax)
  knockoff_selected = names(result$selected)
  
  # Run the multilayer filter
  # MULTILAYER KNOCKOFF FILTER OPTIONS
  fdr.multi = c(q,q)
  knockoff_type = "fixed_equi"
  statistic_type = "group_LSM" 
  FDP_hat_type = "kn+"
  groups.tmp <- get_position(colnames(X))
  all.gen <- unique(groups.tmp)
  groups <- matrix(sapply(groups.tmp, 
                          function(x){which(all.gen==x)}),
                   ncol=1)
  groups <- cbind(1:nrow(groups),groups)
  
  # RUN MULTILAYER KNOCKOFF FILTER
  output = multilayer_knockoff_filter(X, y, groups, fdr.multi,
                                      knockoff_type,
                                      statistic_type, FDP_hat_type)
  knockoff_selected_multi <- colnames(X)[output$S_hat]
  
  # Run the knockoff filter(Second Order).
  knock.gen = function(x) create.second_order(x, method='equi')
  result = knockoff.filter(X, y, fdr=q, knockoffs=knock.gen, statistic=stat.glmnet_lambdasmax)
  #knockoff_selected_2ndO = names(result$selected)
  
  # Run BHq.
  p = ncol(X)
  lm.fit = lm(y ~ X - 1) # no intercept
  p.values = coef(summary(lm.fit))[,4]
  names(p.values) <- substring(names(p.values),2)
  cutoff = max(c(0, which(sort(p.values) <= q * (1:p) / p)))
  bhq_selected = names(which(p.values <= q * cutoff / p))
  
  # Run p-filter
  p_filter = pfilter(p.values, fdr.multi, groups)
  PF =  colnames(X)[which(p_filter==1)]
  # PF Result
  #PF = PF(knockoff_selected, PF(knockoff_selected_2ndO, bhq_selected))
  list(Knockoff = knockoff_selected,MKF = knockoff_selected_multi,
       BHq = bhq_selected, PF = PF)
}
```


```{r Flatten, echo=FALSE}
# Flatten a matrix to a vector with names from concatenating row/column names.
flatten_matrix <- function(M, sep='.') {
  x <- c(M)
  names(x) <- c(outer(rownames(M), colnames(M),
                      function(...) paste(..., sep=sep)))
  x
}

Flatten_X <- function(Drug){
  gene_df <- Drug$gene_df
  muts = c(LETTERS, 'i', 'd')
  pos_start = which(names(gene_df) == 'P1')
  
  X = outer(muts, as.matrix(gene_df[,pos_start:ncol(gene_df)]),
            Vectorize(grepl))
  X = aperm(X, c(2,3,1))
  dimnames(X)[[3]] <- muts
  X = t(apply(X, 1, flatten_matrix))
  mode(X) <- 'numeric'
  return(X)
}

```


To begin with, we have to flatten the matrix of the gene mutations associated with paticular genes.
```{r Flatten_X}
res <- res_PI
X <- Flatten_X(res)
```

After that, we apply Knockoff and BHq procedure to select features.
```{r run_fun}
fdr = 0.20

get_position <- function(x)
  sapply(regmatches(x, regexpr('[[:digit:]]+', x)), as.numeric)

# Define Datasets for PIs
gene_df <- res$gene_df
pos_start <- which(names(gene_df) == 'P1')
tsm_df <- res$tsm_df
Y = gene_df[,4:(pos_start-1)]

# Run Knockoff and BHq algorithm
results = lapply(Y, function(y) knockoff_and_bhq(X, y, fdr))
```

```{r Extract_info_pos, include=FALSE}

comparisons_pos <- lapply(results, function(drug) {
  lapply(drug, function(selected) {
    positions = unique(get_position(selected)) # remove possible duplicates
    discoveries = length(positions)
    false_discoveries = length(setdiff(positions, tsm_df$Position))
    list(true_discoveries = discoveries - false_discoveries,
         false_discoveries = false_discoveries,
         fdp = false_discoveries / max(1, discoveries),
         dp = (discoveries - false_discoveries) / length(tsm_df$Position))
  })
})


plot_data_pos <- NULL
for (drug in names(comparisons_pos)) {
  plot_data <- do.call(cbind, comparisons_pos[[drug]])
  
  plot_data_pos <- rbind(plot_data_pos,
                         data.frame(KF = 
                           as.numeric(plot_data[,'Knockoff']),
                         MKF =
                           as.numeric(plot_data[,'MKF']),
                         BH = as.numeric(plot_data[,'BHq']),
                         PF = as.numeric(plot_data[,'PF']),
                         Criteria = rownames(plot_data),
                         drug = rep(drug,nrow(plot_data))))
}

```


```{r Extract_info_PosMut, include=FALSE}
comparisons_posmut <- lapply(results, function(drug) {
  lapply(drug, function(selected) {
    selected <- unique(selected)
    discoveries = length(selected)
    false_discoveries = length(setdiff(selected, PosMut))
    list(true_discoveries = discoveries - false_discoveries,
         false_discoveries = false_discoveries,
         fdp = false_discoveries / max(1, discoveries),
         dp = (discoveries - false_discoveries) / length(PosMut))
  })
})

plot_data_posmut <- NULL
for (drug in names(comparisons_posmut)) {
  plot_data <- do.call(cbind, comparisons_posmut[[drug]])
  
  plot_data_posmut <- rbind(plot_data_posmut,
                         data.frame(KF = 
                           as.numeric(plot_data[,'Knockoff']),
                         MKF =
                           as.numeric(plot_data[,'MKF']),
                         BH = as.numeric(plot_data[,'BHq']),
                         PF = as.numeric(plot_data[,'PF']),
                         Criteria = rownames(plot_data),
                         drug = rep(drug,nrow(plot_data))))
}

```

```{r plot_data_1}
Criteria <- c("fdp","dp")
plot_data <- function(data, criteria){
  data <- 
    data %>% 
    filter( Criteria %in% criteria) 
  data <- melt(data,  id=c("Criteria","drug","Layer"))
  data <- reshape2::dcast(data, drug+variable+Layer~Criteria,
                          value.var = "value")
  p <- 
    ggplot(data, aes(x=fdp, y =dp, 
                     color = variable))+
    geom_point()+
    geom_vline(aes(xintercept=fdr), linetype = 2)+
    facet_wrap(.~Layer)
}

plot_data.sep <- rbind(plot_data_pos %>% 
                         mutate(Layer="Layer: Position"),
                       plot_data_posmut %>%
                         mutate(Layer="Layer: Position+Direction"))
p.sep <- plot_data(plot_data.sep, Criteria)
p.sep

ggsave("Sep.eps", width = 8, height = 4)
```

```{r multi_task, include=FALSE}
knockoff_and_bhq.combine <- function (X, Y, q) {
  # Combine Y together
  n <- nrow(X)
  p <- ncol(X)
  num.task <- ncol(Y)
  drug.type <- rep(names(Y), each = nrow(X))
  X.new <- matrix(0, ncol = ncol(X)*num.task, nrow = num.task*nrow(X))
  for(i in 1:num.task){
    ind.x <- (1+(i-1)*n) : (i*n )
    ind.y <- (1+(i-1)*p) : (i*p )
    X.new[ind.x, ind.y] <- X
  }
  colnames(X.new) <- rep(colnames(X),times = num.task)
  
  y <- as.numeric(unlist(Y))
    
  # Log-transform the drug resistance measurements.
  y = log(y)
  
  # Remove patients with missing measurements.
  missing = is.na(y)
  y = y[!missing]
  X.new = X.new[!missing,]
  drug.type = drug.type[!missing]
  
  # Remove predictors that appear less than 3 times.
  X.new.ind <- colSums(X.new) >= 3
  X.new = X.new[,X.new.ind]
    
  # Remove duplicate predictors.
  X.new.ind <- colSums(abs(cor(X.new)-1) < 1e-4) == 1
  X.new = X.new[,X.new.ind]
  
  gene.groups.tmp <- colnames(X.new)
  gene.all.gen <- unique(gene.groups.tmp)
  gene.group <- matrix(sapply(gene.groups.tmp,
                               function(x){which(gene.all.gen==x)}),
                        ncol=1)
  # Run the group knockoff.
  
  # Run the multilayer filter
  fdr.multi = c(q,q)
  knockoff_type = "fixed_equi"
  statistic_type = "group_LSM" 
  FDP_hat_type = "kn+"
  
  # group_knockoff_single pos+direction
  
  output.group = multilayer_knockoff_filter(X.new, y, gene.group, c(q),
                                            knockoff_type,
                                            statistic_type, FDP_hat_type)
  knockoff_selected.group <- colnames(X.new)[output.group$S_hat]
  
  # MULTILAYER KNOCKOFF FILTER OPTIONS
  groups.tmp <- get_position(colnames(X.new))
  all.gen <- unique(groups.tmp)
  groups <- matrix(sapply(groups.tmp, 
                          function(x){which(all.gen==x)}),
                   ncol=1)
  groups <- cbind(gene.group,groups)
  
  # RUN MULTILAYER KNOCKOFF FILTER
  output = multilayer_knockoff_filter(X.new, y, groups, fdr.multi,
                                      knockoff_type,
                                      statistic_type, FDP_hat_type)
  knockoff_selected_MKF.group <- colnames(X.new)[output$S_hat]
  
  # Run BHq.
  p = ncol(X.new)
  lm.fit = lm(y ~ X.new - 1) # no intercept
  p.values = coef(summary(lm.fit))[,4]
  names(p.values) <- substring(names(p.values),6)
  cutoff = max(c(0, which(sort(p.values) <= q * (1:p) / p)))
  bhq_selected = names(which(p.values <= q * cutoff / p))
  
  # Run p-filter
  BH.group <- pfilter(p.values, q, gene.group)
  BH.selected <- colnames(X.new)[which(BH.group==1)]
  
  # Run p-filter
  p_filter = pfilter(p.values, fdr.multi, groups)
  PF =  colnames(X.new)[which(p_filter==1)]
  # PF Result
  #PF = PF(knockoff_selected, PF(knockoff_selected_2ndO, bhq_selected))
  list(Knockoff = knockoff_selected.group,
       MKF = knockoff_selected_MKF.group,
       BHq = BH.selected, 
       bhq = bhq_selected,
       PF = PF)
}

```

```{r}
result.combine <- knockoff_and_bhq.combine(X, Y, fdr) 
  
```

```{r}
comparisons_pos.combine <- 
 lapply(result.combine, function(selected) {
    positions = unique(get_position(selected)) # remove possible duplicates
    discoveries = length(positions)
    false_discoveries = length(setdiff(positions, tsm_df$Position))
    list(true_discoveries = discoveries - false_discoveries,
         false_discoveries = false_discoveries,
         fdp = false_discoveries / max(1, discoveries),
         dp = (discoveries - false_discoveries) / length(tsm_df$Position))
  })
```


```{r}
comparisons_posmut.combine <- 
  lapply(result.combine, function(selected) {
    selected <- unique(selected)
    discoveries = length(selected)
    false_discoveries = length(setdiff(selected, PosMut))
    list(true_discoveries = discoveries - false_discoveries,
         false_discoveries = false_discoveries,
         fdp = false_discoveries / max(1, discoveries),
         dp = (discoveries - false_discoveries) / length(PosMut))
  })
```



```{r}
plot_data.tmp <-  do.call(cbind, comparisons_pos.combine)
  
plot_data_pos.combine <- data.frame(
                                 GKF =  
                                   as.numeric(plot_data.tmp[,'Knockoff']),
                                  MKF = as.numeric(plot_data.tmp[,'MKF']),
                                  GBH = as.numeric(plot_data.tmp[,'BHq']),
                                  BH = as.numeric(plot_data.tmp[,'bhq']),
                                  PF = as.numeric(plot_data.tmp[,'PF']),
                                  Criteria = rownames(plot_data.tmp),
                                  drug = rep("PI",nrow(plot_data.tmp)))


plot_data.tmp <-  do.call(cbind, comparisons_posmut.combine)
  
plot_data_posmut.combine <- data.frame(
                                 GKF =  
                                   as.numeric(plot_data.tmp[,'Knockoff']),
                                  MKF = as.numeric(plot_data.tmp[,'MKF']),
                                  GBH = as.numeric(plot_data.tmp[,'BHq']),
                                  BH = as.numeric(plot_data.tmp[,'bhq']),
                                  PF = as.numeric(plot_data.tmp[,'PF']),
                                  Criteria = rownames(plot_data.tmp),
                                  drug = rep("PI",nrow(plot_data.tmp)))
```


```{r}
Criteria <- c("fdp","dp")

plot_data.combine <- rbind(plot_data_pos.combine %>% 
                         mutate(Layer="Layer: Position"),
                       plot_data_posmut.combine %>%
                         mutate(Layer="Layer: Position+Direction"))
p.combine <- plot_data(plot_data.combine, Criteria)
p.combine

ggsave("Combine.eps", width = 8, height = 4)
```

